---
title: "Exploring Refrigerated Truck Dataset"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/JacquelineLincroft/iCloud Drive (Archive) - 2Documents/INSH2102 Bostonography/Assignments/FinalProject")
```

# Import Data:
```{r libraries}
library(ggplot2)
library(tidyverse)
library(rgdal)
library(broom)
library(maps)
library(plyr)
library(dplyr)
library(maptools)
library(RColorBrewer)
library(tigris)
```

```{r import-data}
setwd("/Users/JacquelineLincroft/Documents/INSH2102 Bostonography/Assignments/FinalProject/")
truckdf <- read.table(file="Ref_Trucks.csv", header=TRUE, sep=",")

# to view destination and origin lists: ----------
destination_cities <- unique(truckdf$Destination)
origin_regions <- unique(truckdf$Region)

# separate out the data for boston
bostondf <- truckdf[truckdf$Destination == "BOSTON", ]
bostondf <- bostondf[order(bostondf$Origin), ]
boston_mean <- mean(bostondf$Distance, na.rm=TRUE)

# separate out the data for LA
ladf <- truckdf[truckdf$Destination == "LOS ANGELES", ]
la_mean <- mean(ladf$Distance, na.rm=TRUE)
```


# Basic Visualizations:

```{r basic-visualizations, message=FALSE, warning=FALSE}

# how much data do we have for each of the destination cities?
hist_destination <- ggplot(data=truckdf, aes(x=Destination)) + geom_histogram(stat="count")
hist_destination

# which regions does boston's food come from
hist_boston_region <- ggplot(data=bostondf, aes(x=Region)) + geom_histogram(stat="count")
hist_boston_region

# which regions send food the furthest distances?
density_region<-ggplot(data=truckdf, aes(x=Distance, fill=Region)) + scale_fill_hue(name="Region of Origin", labels=unique(truckdf$Region)) + geom_density()
density_region

# which regions send food the furthest distances (to boston)?
density_boston_region<-ggplot(data=bostondf, aes(x=Distance, fill=Region)) + scale_fill_hue(name="Region of Origin", labels=unique(bostondf$Region)) + geom_density()
density_boston_region

```

From these visualizations, it looks like food from Texas and New York generally travel the shortest distances, and food from Mexico/Arizona/Southeast have a lot of variability in the distance travelled, although this may be due to the specificity/generality of the way these regions are categorized.


# Trying Out Some Geospatial Visualizations:

```{r map-region-names, message=FALSE, warning=FALSE}
# to map regions to states ----------
mapping = data.frame(Region=c("Arizona","California","Floria", "Great Lakes", "Mexico-Arizona", "New York", "Other", "PNW", "Southeast", "Texas", "Mexico-Texas", "Midatlantic", "Indiana", "Mexico-California"), region=c("arizona","california","florida", "michigan", "mexico", "new york", "", "washington", "alabama", "texas", "mexico", "virginia", "indiana", "mexico"))
truckdf <- full_join(truckdf, mapping, by=c("Region"))
```



```{r maps-setup}
# helpful code referenced: https://stackoverflow.com/questions/21863359/how-can-i-resize-maps-to-fill-the-plot-window

# old way of getting map data: ----------
# get usa map from "maps" package
usa_mp <- map("state", fill = TRUE, plot = FALSE)
USA_SP <- map2SpatialPolygons(usa_mp, IDs = usa_mp$names, 
                              proj4string = CRS("+proj=longlat +datum=WGS84"))
USA_DATA <- data.frame(seq_len(length(USA_SP)), row.names = names(USA_SP))
usa_spdf <- SpatialPolygonsDataFrame(USA_SP, data = USA_DATA)
# get mexico map from "maps" package
mexico_mp <- map(database = "world", regions = "Mexico", fill=TRUE, plot=FALSE)
MEX_SP <- map2SpatialPolygons(mexico_mp, IDs = mexico_mp$names, 
                              proj4string = CRS("+proj=longlat +datum=WGS84"))
MEX_DATA <- data.frame(seq_len(length(MEX_SP)), row.names = names(MEX_SP))
mexico_spdf <- SpatialPolygonsDataFrame(MEX_SP, data = MEX_DATA)
# plot them
spplot(usa_spdf, col.regions = "transparent", colorkey = FALSE,
       par.settings = list(axis.line = list(col = "transparent")))
spplot(mexico_spdf, col.regions = "transparent", colorkey = FALSE,
       par.settings = list(axis.line = list(col = "transparent")))

usamexisco_spdf <- merge(usa_spdf, mexico_spdf[1,])
save(truckdf, usa_spdf, mexico_spdf, file="TruckData.RData")
```

```{r counts}
library(plyr)
region_item_count <- count(truckdf, vars=c('Region'))
region_total_agg <- aggregate(truckdf, by = list(truckdf$Region), FUN = count)

ggplot() +
  geom_polygon(data = usa_spdf, aes(fill = state_count, outlier.color=NULL, x = long, y = lat, group = group)) +
  theme_void() + coord_map()

```

```{r choropleth}
# to see color palette options
#display.brewer.all(type="seq")

# make our own color palette
pal <- brewer.pal(8, "GnBu")

# choropleth
spplot(states, col.regions = pal, cuts=input$cuts, colorkey = TRUE,
       par.settings = list(axis.line = list(col = pal)))
```

```{r text-searching}
string <- "CABBAGE"
selected_string <- truckdf[grep(string, truckdf$Commodity), ]
```


```{r merging-data}
# helpful reference: https://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html

# new (better) way of getting map data: ----------
states <- map_data("state")
mexico <- map_data("world", regions="Mexico")

# merge the two maps
merged <- rbind(states, mexico)

# merge the map data with truck data 
merged_data <- full_join(merged, truckdf) # TO FIX: make a smaller modified truckdf with only the info we want
#merged_data <- geo_join(usa_spdf, truckdf, "ID", "State", how="left")
#merged_data <- merge(merged, truckdf, by.x=c("region"), by.y=c("State"))

```

